Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-16 ind_files/05_unification_discussion.tex
Move summary table from chap 5 to chap 6

\justifying
In this thesis, we have seen how stochastic birth-death processes can be used to construct and analyze mechanistic individual-based models for the dynamics of finite populations. In doing so, we have also seen that various well-known equations of evolutionary dynamics can be recovered in the infinite population size limit. In the finite-dimensional case, the infinite population limit corresponds to the equations of population genetics and evolutionary game theory. In the infinite-dimensional case, we instead obtain the equations of quantitative genetics. In both cases, the mean value of the trait in the population changes according to an equation resembling the Price equation. My derivation highlights the natural connections between various equations of population dynamics --- For example, the same procedures that lead to the replicator-mutator equation in the case of discretely varying traits yield Kimura's continuum-of-alleles model in the quantitative case --- and extends these similarities to finite, fluctuating populations (Table \ref{table_summary}).

{\centering\begin{sideways}
    \begin{minipage}{\textheight}
        \resizebox{\textheight}{!}{%
            \setstretch{1.5}
            \begin{tabular}{ %I manually specified the width of each column by trial and error
  |p{\dimexpr.25\linewidth-2\tabcolsep-1.3333\arrayrulewidth}% column 1
  |p{\dimexpr.27\linewidth-2\tabcolsep-1.3333\arrayrulewidth}% column 2
  |p{\dimexpr.33\linewidth-2\tabcolsep-1.3333\arrayrulewidth}% column 3
  |p{\dimexpr.25\linewidth-2\tabcolsep-1.3333\arrayrulewidth}% column 4
  |p{\dimexpr.4\linewidth-2\tabcolsep-1.3333\arrayrulewidth}|% column 5
  }
            \hline
         \centering \textbf{Number of possible distinct types/trait variants ($m$)} & \centering \textbf{State Space} &
\centering \textbf{Model parameters} & \centering \textbf{Mesoscopic description} & \centering\arraybackslash \textbf{Infinite population limit} \\
        \hline
        $m = 1$ \newline (Identical individuals)  & $[0,1,2,3,\ldots]$ \newline (Population size) & Two non-negative functions, $b(n)$ and $d(n)$, describing the birth and death rate of individuals when the population size is $n$ & Univariate Fokker-Planck equation \newline (one-dimensional SDEs) & Dynamics of populations of identical individuals\\ 
        \hline
        $1 < m < \infty$ \newline (Discrete traits) & $[0,1,2,3,\ldots]^{m}$ \newline (Number of individuals of each trait variant) & $2m$ non-negative functions, $b_i(\mathbf{n})$ and $d_i(\mathbf{n})$ (for $1 \leq i \leq m$) describing the birth and death rate of trait variant $i$ when the population is $\mathbf{n}=[n_1,\ldots,n_m]$ & Multivariate Fokker-Planck equation \newline ($m$-dimensional SDEs) &  Evolutionary game theory \newline
        Lotka-Volterra competition \newline Quasispecies equation \newline Price equation (discrete traits) \\
        \hline
        $m = \infty$ \newline (Quantitative traits) & $ \left\{\sum\limits_{i=1}^{n}\delta_{x_i} \ \bigg{|} \ n \in \mathbb{N}, \ x_i \in \mathcal{T} \subseteq \mathbb{R}\right\}$ \newline \newline (Each Dirac mass $\delta_{x_i}$ is an individual with trait value $x_i$ in the trait space $\mathcal{T}$) & Two non-negative functionals $b(x|\nu)$ and $d(x|\nu)$ describing the birth and death rate of trait variant $x$ when the population is the (stochastic) field/function $\nu$ & Functional Fokker-Planck equation/Field theory \newline (SPDEs) & Kimura's continuum-of-alleles model \newline~\cite{sasaki_oligomorphic_2011}'s Oligomorphic Dynamics \newline~\cite{wickman_theoretical_2022}'s Trait Space Equations for intraspecific trait variation \newline Gradient Dynamics \newline Price equation (quantitative traits)\\
        \hline
            \end{tabular}
        }
        %\renewcommand\thetable{1}
        \captionof{table}{Summary of the various birth-death processes studied in this thesis}
        \label{table_summary}
    \end{minipage}
\end{sideways}\par}
\clearpage

>>>>

So far, we have seen a lot of relatively abstract formalism for describing evolution in finite populations. Some concrete and hopefully illustrative examples are also presented in Appendix \ref{App_examples}. Have we gained anything from this (re)formulation? I show below that the formalism developed in part \ref{part_theory} naturally yields some `fundamental equations' for evolutionary population dynamics that are very general, help us clearly understand how evolution operates in finite fluctuating populations, and recover well-known results in the infinite population limit, thus showing that I have generalized these infinite population results.

### 2023-03-16 ind_files/06_nontechnical_discussion.tex
Move summary table from chap 5 to chap 6

>>>>

\justifying
In this thesis, we have seen how stochastic birth-death processes can be used to construct and analyze mechanistic individual-based models for the dynamics of finite populations. In doing so, we have also seen that various well-known equations of evolutionary dynamics can be recovered in the infinite population size limit. In the finite-dimensional case, the infinite population limit corresponds to the equations of population genetics and evolutionary game theory. In the infinite-dimensional case, we instead obtain the equations of quantitative genetics. In both cases, the mean value of the trait in the population changes according to an equation resembling the Price equation. My derivation highlights the natural connections between various equations of population dynamics --- For example, the same procedures that lead to the replicator-mutator equation in the case of discretely varying traits yield Kimura's continuum-of-alleles model in the quantitative case --- and extends these similarities to finite, fluctuating populations (Table \ref{table_summary}).

{\centering\begin{sideways}
		\begin{minipage}{\textheight}
			\resizebox{\textheight}{!}{%
				\setstretch{1.5}
				\begin{tabular}{ %I manually specified the width of each column by trial and error
						|p{\dimexpr.25\linewidth-2\tabcolsep-1.3333\arrayrulewidth}% column 1
						|p{\dimexpr.27\linewidth-2\tabcolsep-1.3333\arrayrulewidth}% column 2
						|p{\dimexpr.33\linewidth-2\tabcolsep-1.3333\arrayrulewidth}% column 3
						|p{\dimexpr.25\linewidth-2\tabcolsep-1.3333\arrayrulewidth}% column 4
						|p{\dimexpr.4\linewidth-2\tabcolsep-1.3333\arrayrulewidth}|% column 5
					}
					\hline
					\centering \textbf{Number of possible distinct types/trait variants ($m$)} & \centering \textbf{State Space} &
					\centering \textbf{Model parameters} & \centering \textbf{Mesoscopic description} & \centering\arraybackslash \textbf{Infinite population limit} \\
					\hline
					$m = 1$ \newline (Identical individuals)  & $[0,1,2,3,\ldots]$ \newline (Population size) & Two non-negative functions, $b(n)$ and $d(n)$, describing the birth and death rate of individuals when the population size is $n$ & Univariate Fokker-Planck equation \newline (one-dimensional SDEs) & Dynamics of populations of identical individuals\\ 
					\hline
					$1 < m < \infty$ \newline (Discrete traits) & $[0,1,2,3,\ldots]^{m}$ \newline (Number of individuals of each trait variant) & $2m$ non-negative functions, $b_i(\mathbf{n})$ and $d_i(\mathbf{n})$ (for $1 \leq i \leq m$) describing the birth and death rate of trait variant $i$ when the population is $\mathbf{n}=[n_1,\ldots,n_m]$ & Multivariate Fokker-Planck equation \newline ($m$-dimensional SDEs) &  Evolutionary game theory \newline
					Lotka-Volterra competition \newline Quasispecies equation \newline Price equation (discrete traits) \\
					\hline
					$m = \infty$ \newline (Quantitative traits) & $ \left\{\sum\limits_{i=1}^{n}\delta_{x_i} \ \bigg{|} \ n \in \mathbb{N}, \ x_i \in \mathcal{T} \subseteq \mathbb{R}\right\}$ \newline \newline (Each Dirac mass $\delta_{x_i}$ is an individual with trait value $x_i$ in the trait space $\mathcal{T}$) & Two non-negative functionals $b(x|\nu)$ and $d(x|\nu)$ describing the birth and death rate of trait variant $x$ when the population is the (stochastic) field/function $\nu$ & Functional Fokker-Planck equation/Field theory \newline (SPDEs) & Kimura's continuum-of-alleles model \newline~\cite{sasaki_oligomorphic_2011}'s Oligomorphic Dynamics \newline~\cite{wickman_theoretical_2022}'s Trait Space Equations for intraspecific trait variation \newline Gradient Dynamics \newline Price equation (quantitative traits)\\
					\hline
				\end{tabular}
			}
			%\renewcommand\thetable{1}
			\captionof{table}{Summary of the various birth-death processes studied in this thesis}
			\label{table_summary}
		\end{minipage}
	\end{sideways}\par}
\clearpage
